\textsf{Are large language models superhuman chemists?}

### 2024-10-16 src/tex/ms.tex
implement some human statistics
reading ease

%To contextualize the scores, we also surveyed \variable{output/number_experts.txt} experts in chemistry on a subset of the benchmark corpus to be able to compare the performance of current frontier models with (human) chemists of different specializations.
Our results indicate that current frontier models perform \enquote{superhuman} on some aspects of chemistry but, in many cases, including safety-related ones, might be very misleading.
We find that there are still numerous limitations for current models to overcome, such that they can be directly applied in autonomous systems for chemists.
Moreover, we conclude that carefully created broad benchmarks represent an important stepping stone for progress in this field.  %and underline the importance of careful evaluation of the capabilities of \glspl{llm} in the chemical sciences.

@@

% \begin{figure}[!h]
%     \centering
%     \includegraphics{figures/human_subset_performance.pdf}
%     \caption{\textbf{Performance of models and humans on the \enquote{tiny} subset of the  \chembench corpus.} The figure shows the percentage of questions that the models answered completely correctly. We use horizontal bars to indicate the performance of various models and highlight statistics of the human performance. 
%     Since the humans did not answer all the questions, this plot is based on the subset of questions that most humans answered.
%     The evaluation we use here is very strict as it only considers a question answered completely correctly or completely incorrectly, partially correct answers are also considered incorrect.
%     \Cref{fig:barplot_all_correct_all_questions} provides an overview of the performance of various models on the entire corpus.
%     Systems with \enquote{ReAct} in the name are tool augmented, i.e., they can call external tools such as web search or Python code executors to better answer the questions.
%     However, we limit those systems to a maximum of ten calls to the \gls{llm}. This constraint led the systems to often not find the correct answer within the specified number of calls.
%     In this case, we consider the answer as incorrect. 
%     }
%     \label{fig:human_vs_models_bar}
%     \script{plot_overview_performance_plot.py}
% \end{figure}

@@

\paragraph{Performance per topic} To obtain a more detailed understanding of the performance of the models, we also analyzed the performance of the models in different subfields of the chemical sciences.
For this analysis, we defined a set of topics (see \Cref{sec:meth-topic}). We classified all questions in the \chembench corpus into these topics based on hand-crafted rules and classifier models operating on the question texts.
We then computed the percentage of questions the models or humans answered correctly for each topic.
% \begin{figure}[!h]
%     \centering
%     \includegraphics{figures/all_questions_models_completely_correct_radar_human.pdf}
%     \caption{\textbf{Performance of the models and humans on the different topics of the \enquote{tiny} subset of the \chembench corpus.} The radar plot shows the performance of the models and humans on the different topics of \enquote{tiny} subset of the \chembench corpus. The performance is measured as the fraction of questions that were answered completely correctly by the models.
%     The best score for every dimension is one (all questions answered correctly), and the worst is zero (no question answered completely correctly). A larger colored area indicates a better performance.
%     This figure shows the performance on the subset of questions that were answered by humans. The performance of models on the entire corpus is shown in \Cref{fig:performance_per_topic}.
%     }
%     \label{fig:all_questions_models_completely_correct_radar_human}
%     \script{analyze_model_reports.py}
% \end{figure}

@@

% \begin{figure}[!h]
%     \centering
%     \includegraphics{figures/confidence_vs_performance_overall.pdf}
%     \caption{\textbf{Relationship between confidence of the model in the answer and correctness.} For this analysis, we used verbalized confidence estimates from the model. We prompted the models to return a confidence score on an ordinal scale to obtain those estimates. We then plot the correctness of the answers (which is calculated as the mean value of all answers being either completely correct (1) or completely incorrect (0)) against the confidence score.
%         The stripplots show the individual data points, a strong density indicating a higher number of points. The lines show the mean, and the error bars show the standard error of the mean. The figure shows that the reliability of the confidence estimates varies widely across models. While the ones for GPT-4 and Claude 2 are not particularly reliable, the ones for Claude 3 follow the expected trend better.}
%     \label{fig:confidence_vs_performance}
%     \script{joint_analysis_confidence_performance.py}
% \end{figure}

>>>>

To contextualize the scores, we also surveyed \variable{output/number_experts.txt} experts in chemistry on a subset of the benchmark corpus to be able to compare the performance of current frontier models with (human) chemists of different specializations.
Our results indicate that current frontier models perform \enquote{superhuman} on some aspects of chemistry but, in many cases, including safety-related ones, might be very misleading.
We find that there are still numerous limitations for current models to overcome, such that they can be directly applied in autonomous systems for chemists.
Moreover, we conclude that carefully created broad benchmarks represent an important stepping stone for progress in this field.  %and underline the importance of careful evaluation of the capabilities of \glspl{llm} in the chemical sciences.

@@

\begin{figure}[!h]
    \centering
    \includegraphics{figures/human_subset_performance.pdf}
    \caption{\textbf{Performance of models and humans on the \enquote{tiny} subset of the  \chembench corpus.} The figure shows the percentage of questions that the models answered correctly. We use horizontal bars to indicate the performance of various models and highlight statistics of the human performance. 
    Since the humans did not answer all the questions, this plot is based on the subset of questions that most humans answered.
    The evaluation we use here is very strict as it only considers a question answered correctly or incorrectly, partially correct answers are also considered incorrect.
    \Cref{fig:barplot_all_correct_all_questions} provides an overview of the performance of various models on the entire corpus.
    Systems with \enquote{ReAct} in the name are tool augmented, i.e., they can call external tools such as web search or Python code executors to better answer the questions.
    However, we limit those systems to a maximum of ten calls to the \gls{llm}. This constraint led the systems to often not find the correct answer within the specified number of calls.
    In this case, we consider the answer as incorrect. 
    }
    \label{fig:human_vs_models_bar}
    \script{plot_overview_performance_plot.py}
\end{figure}

@@

\paragraph{Performance per topic} To obtain a more detailed understanding of the performance of the models, we also analyzed the performance of the models in different subfields of the chemical sciences.
For this analysis, we defined a set of topics (see \Cref{sec:meth-topic}). We classified all questions in the \chembench corpus into these topics based on hand-crafted rules and classifier models operating on the question texts.
We then computed the percentage of questions the models or humans answered correctly for each topic.
% \begin{figure}[!h]
%     \centering
%     \includegraphics{figures/all_questions_models_completely_correct_radar_human.pdf}
%     \caption{\textbf{Performance of the models and humans on the different topics of the \enquote{tiny} subset of the \chembench corpus.} The radar plot shows the performance of the models and humans on the different topics of \enquote{tiny} subset of the \chembench corpus. The performance is measured as the fraction of questions that were answered completely correctly by the models.
%     The best score for every dimension is one (all questions answered correctly), and the worst is zero (no question answered correctly). A larger colored area indicates a better performance.
%     This figure shows the performance on the subset of questions that were answered by humans. The performance of models on the entire corpus is shown in \Cref{fig:performance_per_topic}.
%     }
%     \label{fig:all_questions_models_completely_correct_radar_human}
%     \script{analyze_model_reports.py}
% \end{figure}

@@

% \begin{figure}[!h]
%     \centering
%     \includegraphics{figures/confidence_vs_performance_overall.pdf}
%     \caption{\textbf{Relationship between confidence of the model in the answer and correctness.} For this analysis, we used verbalized confidence estimates from the model. We prompted the models to return a confidence score on an ordinal scale to obtain those estimates. We then plot the correctness of the answers (which is calculated as the mean value of all answers being either correct (1) or incorrect (0)) against the confidence score.
%         The stripplots show the individual data points, a strong density indicating a higher number of points. The lines show the mean, and the error bars show the standard error of the mean. The figure shows that the reliability of the confidence estimates varies widely across models. While the ones for GPT-4 and Claude 2 are not particularly reliable, the ones for Claude 3 follow the expected trend better.}
%     \label{fig:confidence_vs_performance}
%     \script{joint_analysis_confidence_performance.py}
% \end{figure}
